---
title: "Final Project"
author: "Safiya Alavi and Maya Sinha"
date: "2022-10-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview of Data Set

<br> 

We are analyzing the quality of wine based on 11 predictors: fixed acidity, volatile acidity, citric acid, residual sugar, chlorides, free sulfur dioxide, total sulfur dioxide, density, pH, sulfates, alcohol. Quality, in this instance, is defined as an indicator of its craftsmanship and, thus, desirability which can be used, for example, in pricing. Quality does not necessarily indicate if a wine has gone bad. We will use two data sets -- one for white wine and the other for red -- to create two respective machine learning models. There are 4899 observations in the white wine data set and 1600 observations in the red wine data set. 
>>>>>>> 4107e5d (.)

## Overview of Research Questions 

<br>

We are interested in conducting analysis on both the white and red wine to assess the quality of wine in new wines in production so that they may be accurately priced and marketed.  

## Proposed Project Timeline 

<br>

Week 2: Data cleaning<br>
Week 3-5: Exploratory Data Analysis<br>
Week 6-8: Model Building<br>
Week 9: Conclusion and final touches<br>

## Any questions or concerns

<br>


# Loading Data and Packages 
This project uses data on the white and red wine, which records information about the details of the wine. 

```{r load packages and data}
# load packages
library(tidyverse)
library(lubridate)
library(tidymodels)
library(dplyr)

white_train <- read.csv("/Users/safiyaalavi/Desktop/PSTAT 131/131FinalProject/WineDataset/winequality-white.csv", sep = ";")

red_train <- read.csv("/Users/safiyaalavi/Desktop/PSTAT 131/131FinalProject/WineDataset/winequality-red.csv", sep = ";")

# Make Column Names Clean 
colnames(white_train) <- c("Fixed Acidity", "Volatile Acidity","Citric Acid", "Residual Sugar", "Chlorides", "Free Sulfur Dioxide", "Total Sulfur Dioxide", "Density", "pH", "Sulfates", "Alcohol", "Quality")
colnames(red_train) <- c("Fixed Acidity", "Volatile Acidity","Citric Acid", "Residual Sugar", "Chlorides", "Free Sulfur Dioxide", "Total Sulfur Dioxide", "Density", "pH", "Sulfates", "Alcohol", "Quality")

```


# Data Cleaning 

# Data Splitting 

=======
Does the age of the wine also play a role in this? We should also find out if that data we are using is taken from newly produced wine or is taken by measuring nonorganized cases of wine.
 

## Data Cleaning
## EDA
## Model Fitting
>>>>>>> 4107e5d (.)