---
title: "Final Project"
author: "Safiya Alavi and Maya Sinha"
date: "2022-10-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Guidelines
#### Code
Uses the tidyverse and tidymodels for fitting
models, etc. includes comments as necessary.
Uses code folding so huge blocks of code
aren't displayed when knitting.

#### Narration
Includes introduction and conclusion sections.
Good flow of text and narration throughout; at
least some description/explanation for every
piece of code/results. Easily readable. Makes
an effort to explain/make sense to general
audience. Written like a paper. Divided into
reasonable sections.

#### Exploratory Data Analysis
Conducts some EDA. Explores outcome
variable distribution. Assesses missing data
patterns. Creates at least 3-5 plots or tables to
explore relationships among variables.

#### Organization and Structure
Submits a link to a GitHub repository
containing your project, with .Rmd and knitted
.html or .pdf. Includes at least one subfolder
for data (with subdirectories). Includes
codebook and data citation.

#### Feature Extraction
Creates appropriate models. Handles
categorical predictors reasonably (with
dummy/one-hot encoding, PCA, etc.). Creates
interactions if necessary. Justifies any variable
transformations.

#### Data Splitting
Divides data into training and test sets with a
reasonable proportion. Uses stratified
sampling. Uses cross-validation to fold
training set. (Ideally, stratified CV with
repeats.)

#### Model Fitting
Fits at least 4 model classes (random forest,
KNN, boosted tree, lasso or ridge regression or
logistic regression, SVM, neural network).
Tunes models across resamples. Fits optimal
tuned model to training and test sets.

## Overview of Data Set

<br> 

We are analyzing the quality of wine based on 11 predictors: fixed acidity, volatile acidity, citric acid, residual sugar, chlorides, free sulfur dioxide, total sulfur dioxide, density, pH, sulfates, alcohol. Quality, in this instance, is defined as an indicator of its craftsmanship and, thus, desirability which can be used, for example, in pricing. Quality does not necessarily indicate if a wine has gone bad. We will use two data sets -- one for white wine and the other for red -- to create two respective machine learning models. There are 4899 observations in the white wine data set and 1600 observations in the red wine data set. 
>>>>>>> 4107e5d (.)

## Overview of Research Questions 

<br>

We are interested in conducting analysis on both the white and red wine to assess the quality of wine in new wines in production so that they may be accurately priced and marketed.  

## Proposed Project Timeline 

<br>

Week 2: Data cleaning<br>
Week 3-5: Exploratory Data Analysis<br>
Week 6-8: Model Building<br>
Week 9: Conclusion and final touches<br>

## Any questions or concerns

<br>


# Loading Data and Packages 
This project uses data on the white and red wine, which records information about the details of the wine. 

```{r load packages and data}
# load packages
library(tidyverse)
library(lubridate)
library(tidymodels)

white_train <- read.csv("/Users/safiyaalavi/Desktop/PSTAT 131/131FinalProject/WineDataset/winequality-white.csv", sep = ";")

red_train <- read.csv("/Users/safiyaalavi/Desktop/PSTAT 131/131FinalProject/WineDataset/winequality-red.csv", sep = ";")

```


# Data Cleaning 

# Data Splitting 

=======
Does the age of the wine also play a role in this? We should also find out if that data we are using is taken from newly produced wine or is taken by measuring nonorganized cases of wine.
 

## Data Cleaning
## EDA
## Model Fitting
>>>>>>> 4107e5d (.)